---
title: "Project 2: A sentiment analysis on different cities in Vietnam"
output:
  pdf_document: default
  html_document: default
date: "2024-05-21"
---

```{r}
# Load necessary libraries
library(tidyverse)  # For data manipulation
library(readr)      # For reading CSV files
library(lubridate)  # For handling dates
library(leaflet)    # For creating maps
library(syuzhet)
library(reticulate)
library(vader)
library(textTinyR)
library(textdata)
library(dplyr)
library(readr)
library(shiny)
library(leaflet)
library(sf)


```
# Read and prepare the data 

```{r}
# Load the dataset
df <- read_csv("data/reddit_tourism_headlines_with_sentiment.csv")

# View the first few rows of the dataset
head(df)

# Summary of the dataset to understand its structure
summary(df)
```

# Preprocess the Data

```{r}
# Convert the Date column to Date type
df$Date <- as.Date(df$Date, format="%Y-%m-%d")

# Aggregate sentiment scores by city
city_sentiment <- df %>%
  group_by(City) %>%
  summarise(
    compound = mean(compound, na.rm = TRUE),
    positive = mean(positive, na.rm = TRUE),
    neutral = mean(neutral, na.rm = TRUE),
    negative = mean(negative, na.rm = TRUE)
  )

# Print the aggregated sentiment scores
print(city_sentiment)

```

#Create Bar Plot that Represents The Distribution of Sentiment Scores
```{r}

sentiment_data <- df %>%
  mutate(sentiment_category = case_when(
    compound > 0 ~ "Positive",
    compound < 0 ~ "Negative",
    TRUE ~ "Neutral"
  ))

sentiment_summary <- sentiment_data %>%
  count(sentiment_category)

ggplot(sentiment_summary, aes(x = sentiment_category, y = n, fill = sentiment_category)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Positive" = "lightgreen", "Negative" = "darkred", "Neutral" = "gray"), name = "Sentiment Category") +
  theme_minimal() +
  labs(title = "Sentiment Analysis Distribution",
       x = "Sentiment Category",
       y = "Count")


```
#Create Bar Plot that Represents The Distribution of Sentiment Scores in Different Cities
```{r}
sentiment_summary <- sentiment_data %>%
  group_by(City, sentiment_category) %>%
  summarise(count = n(), .groups = 'drop')


ggplot(sentiment_summary, aes(x = City, y = count, fill = sentiment_category)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("Positive" = "lightgreen", "Negative" = "darkred", "Neutral" = "gray"), name = "Sentiment Category") +
  theme_minimal() +
  labs(title = "Sentiment Analysis Distribution Across Cities",
       x = "City",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#Create Time Series Plot that that represent the counts of each sentiment category over time for different cities:
```{r}
df$Date <- as.Date(df$Date)

sentiment_summary <- sentiment_data %>%
  group_by(City, Date, sentiment_category) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(sentiment_summary, aes(x = Date, y = count, color = sentiment_category)) +
  geom_line() +
  facet_wrap(~ City, scales = "free_y") +
  scale_color_manual(values = c("Positive" = "lightgreen", "Negative" = "darkred", "Neutral" = "gray"), name = "Sentiment Category") +
  theme_minimal() +
  labs(title = "Sentiment Analysis Over Time Across Cities",
       x = "Date",
       y = "Count")


```


#Create Interactive Dashboard
```{r}
library(shiny)
library(ggplot2)
library(dplyr)


ui <- fluidPage(
  titlePanel("Vietnam Sentiment Analysis Dashboard"),
  sidebarLayout(
    sidebarPanel(
      dateRangeInput("date_range", "Select Date Range:",
                     start = min(sentiment_data$Date),
                     end = max(sentiment_data$Date),
                     format = "yyyy-mm-dd"),
      selectInput("city", "Select City:",
                  choices = unique(sentiment_data$City),
                  selected = unique(sentiment_data$City)[1])
    ),
    mainPanel(
      plotOutput("sentiment_plot")
    )
  )
)

server <- function(input, output) {
  filtered_data <- reactive({
    sentiment_data %>%
      filter(Date >= input$date_range[1] & Date <= input$date_range[2],
             City == input$city) %>%
      group_by(Date, sentiment_category) %>%
      summarise(count = n(), .groups = 'drop')
  })

  output$sentiment_plot <- renderPlot({
    ggplot(filtered_data(), aes(x = Date, y = count, color = sentiment_category)) +
      geom_line() +
      scale_color_manual(values = c("Positive" = "lightgreen", "Negative" = "darkred", "Neutral" = "gray"), name = "Sentiment Category") +
      theme_minimal() +
      labs(title = paste("Sentiment Analysis in", input$city, "Over Time"),
           x = "Date",
           y = "Count")
  })
}

shinyApp(ui, server)



```
# Create Interactive Choropleth Map

```{r}


# Load necessary libraries
library(sf)
library(ggplot2)
library(dplyr)
library(viridis)

# Define the path to the shapefile (adjust the path as necessary)
shapefile_path <- "shapefile/gadm41_VNM_1.shp"

# Load the shapefile
vietnam_shapefile <- st_read(shapefile_path)

# Check if the shapefile has loaded correctly
print(vietnam_shapefile)

# Create a dataframe with city coordinates
city_coords <- data.frame(
  City = c("Hanoi", "Saigon", "Da Nang", "Sapa", "Nha Trang", "Hoi An", "Hue", "Phu Quoc", "Vung Tau", "Da Lat"),
  Latitude = c(21.028511, 10.823099, 16.054407, 22.337400, 12.238791, 15.880058, 16.463713, 10.289813, 10.346101, 11.940419),
  Longitude = c(105.804817, 106.629664, 108.202167, 103.841179, 109.196749, 108.338046, 107.595169, 103.984018, 107.084271, 108.458313)
)

# Calculate the average sentiment score for each city
average_sentiment <- sentiment_data %>%
  group_by(City) %>%
  summarise(mean_sentiment = mean(compound, na.rm = TRUE))

# Merge with city coordinates
city_coords_sentiment <- city_coords %>%
  left_join(average_sentiment, by = "City")

# Convert to spatial dataframe
city_coords_sentiment_sf <- st_as_sf(city_coords_sentiment, coords = c("Longitude", "Latitude"), crs = 4326)

# Plot the choropleth map with city names and adjusted x and y axis
ggplot() +
  geom_sf(data = vietnam_shapefile) +
  geom_sf(data = city_coords_sentiment_sf, aes(color = mean_sentiment), size = 3) +
  geom_text(data = city_coords, aes(x = Longitude, y = Latitude, label = City), size = 2, vjust = -1) +  
  scale_color_viridis_c(option = "C", na.value = "lightgrey") +
  theme_minimal() +
  theme(axis.title.y = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10)),
        axis.text.y = element_text(size = 6),  
        axis.text.x = element_text(size = 6)) + 
  labs(title = "Sentiment Map of Tourism Headlines in Vietnam",
       color = "Average Sentiment Score",
       x = "Longitude",
       y = "Latitude")


```
```{r}
# Aggregate sentiment scores by city and date
city_sentiment_time <- df %>%
  group_by(City, Date) %>%
  summarise(
    compound = mean(compound, na.rm = TRUE),
    positive = mean(positive, na.rm = TRUE),
    neutral = mean(neutral, na.rm = TRUE),
    negative = mean(negative, na.rm = TRUE),
    .groups = 'drop'
  )

```

```{r}
#library(gganimate)  # For animation
#library(transformr)  # For smoother animations

# Create the animated choropleth map
#p <- ggplot() +
 # geom_sf(data = vietnam_shapefile) +
  #geom_sf(data = city_coords_sentiment_time_sf, aes(color = compound), size = 3) +
  #geom_text(data = city_coords, aes(x = Longitude, y = Latitude, label = City), size = 2, vjust = -1) +
  #scale_color_viridis_c(option = "C", na.value = "lightgrey") +
  #theme_minimal() +
  #theme(axis.title.y = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10)),
  #      axis.text.y = element_text(size = 6),
   #     axis.text.x = element_text(size = 6)) +
  #labs(title = "Sentiment Map of Tourism Headlines in Vietnam",
    #   color = "Average Sentiment Score",
     #  x = "Longitude",
      # y = "Latitude",
       #subtitle = 'Date: {frame_time}') + 
#  transition_time(Date) +
 # ease_aes('linear')

#animate(p, nframes = 100, fps = 5, width = 800, height = 600, renderer = gifski_renderer("animated_map.gif"))

```



```{r}
# Merge with city coordinates
city_coords_sentiment_time <- city_coords %>%
  left_join(city_sentiment_time, by = "City")

# Convert to spatial dataframe
city_coords_sentiment_time_sf <- st_as_sf(city_coords_sentiment_time, coords = c("Longitude", "Latitude"), crs = 4326)

ui <- fluidPage(
  titlePanel("Sentiment Analysis of Tourism Headlines in Vietnam"),
  sidebarLayout(
    sidebarPanel(
      dateRangeInput("dateRange",
                     "Select Date Range:",
                     start = min(city_coords_sentiment_time$Date, na.rm = TRUE),
                     end = max(city_coords_sentiment_time$Date, na.rm = TRUE),
                     min = min(city_coords_sentiment_time$Date, na.rm = TRUE),
                     max = max(city_coords_sentiment_time$Date, na.rm = TRUE)),
      selectInput("city", "Select City:", choices = unique(city_coords_sentiment_time$City), selected = "Hanoi")
    ),
    mainPanel(
      leafletOutput("map")
    )
  )
)

# Define server logic
server <- function(input, output) {
  filtered_data <- reactive({
    city_coords_sentiment_time %>%
      filter(City == input$city, Date >= input$dateRange[1], Date <= input$dateRange[2])
  })
  
  output$map <- renderLeaflet({
    leaflet() %>%
      addTiles() %>%
      addPolygons(data = vietnam_shapefile, fillColor = "lightgrey", color = "black", weight = 1,
                  highlightOptions = highlightOptions(weight = 3, color = "blue", bringToFront = TRUE)) %>%
      setView(lng = 106.629664, lat = 16.054407, zoom = 6) %>%
      addCircleMarkers(
        data = filtered_data(),
        lat = ~Latitude,
        lng = ~Longitude,
        color = ~colorNumeric("viridis", domain = filtered_data()$compound)(compound),
        fillOpacity = 0.7,
        radius = 5,
        popup = ~paste(City, "<br>", "Date:", Date, "<br>", "Sentiment Score:", round(compound, 2))
      ) %>%
      addLegend(
        position = "bottomright",
        pal = colorNumeric("viridis", domain = city_coords_sentiment_time$compound),
        values = city_coords_sentiment_time$compound,
        title = "Sentiment Score",
        opacity = 1
      )
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```

```{r}
# Merge with city coordinates
city_coords_sentiment_time <- city_coords %>%
  left_join(city_sentiment_time, by = "City")

# Convert to spatial dataframe
city_coords_sentiment_time_sf <- st_as_sf(city_coords_sentiment_time, coords = c("Longitude", "Latitude"), crs = 4326)

ui <- fluidPage(
  titlePanel("Sentiment Analysis of Tourism Headlines in Vietnam"),
  sidebarLayout(
    sidebarPanel(
      dateRangeInput("dateRange",
                     "Select Date Range:",
                     start = min(city_coords_sentiment_time$Date, na.rm = TRUE),
                     end = max(city_coords_sentiment_time$Date, na.rm = TRUE),
                     min = min(city_coords_sentiment_time$Date, na.rm = TRUE),
                     max = max(city_coords_sentiment_time$Date, na.rm = TRUE)),
      selectInput("city", "Select City:", choices = unique(city_coords_sentiment_time$City), selected = "Hanoi"),
      sliderInput("sentimentRange",
                  "Select Sentiment Score Range:",
                  min = min(city_coords_sentiment_time$compound, na.rm = TRUE),
                  max = max(city_coords_sentiment_time$compound, na.rm = TRUE),
                  value = c(min(city_coords_sentiment_time$compound, na.rm = TRUE), max(city_coords_sentiment_time$compound, na.rm = TRUE))
      )
    ),
    mainPanel(
      leafletOutput("map"),
      textOutput("meanSentiment")
    )
  )
)

# Define server logic
server <- function(input, output) {
  filtered_data <- reactive({
    city_coords_sentiment_time %>%
      filter(
             Date >= input$dateRange[1],
             Date <= input$dateRange[2],
             compound >= input$sentimentRange[1],
             compound <= input$sentimentRange[2])
  })
  
  mean_sentiment <- reactive({
    filtered_data() %>%
      summarise(mean_compound = mean(compound, na.rm = TRUE)) %>%
      pull(mean_compound)
  })
  
  output$map <- renderLeaflet({
    leaflet() %>%
      addTiles() %>%
      addPolygons(data = vietnam_shapefile, fillColor = "lightgrey", color = "black", weight = 1,
                  highlightOptions = highlightOptions(weight = 3, color = "blue", bringToFront = TRUE)) %>%
      setView(lng = 106.629664, lat = 16.054407, zoom = 6) %>%
      addCircleMarkers(
        data = filtered_data(),
        lat = ~Latitude,
        lng = ~Longitude,
        color = ~colorNumeric("viridis", domain = filtered_data()$compound)(compound),
        fillOpacity = 0.7,
        radius = 5,
        popup = ~paste(City, "<br>", "Date:", Date, "<br>", "Sentiment Score:", round(compound, 2))
      ) %>%
      addLegend(
        position = "bottomright",
        pal = colorNumeric("viridis", domain = city_coords_sentiment_time$compound),
        values = city_coords_sentiment_time$compound,
        title = "Sentiment Score",
        opacity = 1
      )
  })
  
  output$meanSentiment <- renderText({
    mean_value <- mean_sentiment()
    paste("Mean Sentiment Score for Selected Range: ", round(mean_value, 2))
  })
}

# Run the application
shinyApp(ui = ui, server = server)

````
